---
title: "European Energy"
author: "Alex Norman"
date: "04/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My first TidyTuesday attempt!!

OK - let's read the data in using the `tidytuesdayR` package.

```{r message=FALSE, echo=FALSE}
library(tidytuesdayR)

euro_energy <- tidytuesdayR::tt_load(2020, week = 32)
```

This brings a list of two datasets into the global environment.

## Types of dataviz this dataset

* Map with percent of renewable displayed by colour fill
* Map with percent of renewable displayed by colour fill levels 
* Barplot with percent of type displayed by colour fill
* Georgios Karamanis' 2020-Wk8 fan-like graph https://github.com/gkaramanis/tidytuesday/tree/master/2020-week08
    * A variant of this could be with the top 5 and bottom 5, coloured grey and green respectively
* Georgios Karamanis' 2020-Wk7 filled hexes https://github.com/gkaramanis/tidytuesday/tree/master/2020-week07
* 

It seems that for many of these I need percentages, so they need to be created first. The `energy_types` dataframe has the type information needed. I am specifically interested in renewable energy; that is, as opposed to clean energy, which creates little to no greenhouse gas emissions, I want to look at energy that comes from renewable sources, such as sunlight, wind, water, and geothermal heat. So, the goal is to get a data frame for 2018 that summarises energy production by three factors - conventional thermal, nuclear, and renewable.

```{r message=FALSE, echo=FALSE}
library(tidyverse)
library(janitor)

# Narrow the data down to 2018 country observations of energy type
latest_energy <- euro_energy$energy_types %>%
  select(-level) %>%
  mutate(country_name = case_when(country == "EL" ~ "Greece", country == "UK" ~ "United Kingdom", country == "BA" ~ "Bosnia and Herzegovina", TRUE ~ country_name)) %>%
  group_by(country) %>%
  pivot_longer(cols = c(`2016`,`2017`,`2018`), names_to = "year") %>%
  pivot_wider(names_from = type, values_from = value) %>%
  filter(year == 2018) %>%
  janitor::clean_names() %>%
  rowwise() %>%
  mutate(total = sum(conventional_thermal, nuclear, hydro, wind, solar, geothermal, other, pumped_hydro_power),
         across(where(is.numeric), ~ round(100 * . / total, 1))) %>%
  ungroup() %>%
  pivot_longer(names_to = "energy", values_to = "value", 
               c(conventional_thermal, nuclear, hydro, wind, solar, geothermal, other, pumped_hydro_power)) %>%
  mutate(energy = case_when(energy == "conventional_thermal" ~ "Conventional Thermal",
                            energy == "nuclear" ~ "Nuclear",
                            TRUE ~ "Renewable"),
         energy = factor(energy, 
                         levels = c("Conventional Thermal", "Nuclear", 
                                    "Renewable"))) %>%
  group_by(country_name, energy) %>%
  summarise(percent = sum(value)) %>%
  #pivot_wider(names_from = energy, values_from = percent) %>%
  #arrange(desc(Renewable)) %>%
  ungroup()

```
Now to make a plot! First, to get the top and bottom 5 countries by renewable energy production.

```{r plot, message=FALSE, echo=FALSE}
top <- latest_energy %>%
  filter(energy == "Renewable") %>%
  slice_max(n = 10, order_by = percent)

bottom <- latest_energy %>%
  filter(energy == "Renewable") %>%
  slice_min(n = 5, order_by = percent)

top_bot <- top %>%
  #bind_rows(top, bottom) %>%
  arrange(percent) %>%
  mutate(n = -4:5) %>%
  rowwise() %>%
  mutate(
    x = list(c(-10, 0, 0, -10)),
    y = list(c(n*4 - 1.4, n*2 - 0.7, n*2 + 0.7, n*4 + 1.4))
  ) %>% 
  unnest(cols = c(x, y))
  
library(extrafont)
font_import()
ggplot(top_bot) +
  geom_rect(aes(xmin = -60, ymin = n*4 - 1.4,
                xmax = -10, ymax = n*4 + 1.4), fill = "#548b54", color = NA) +
  geom_polygon(aes(x, y, group = n), fill = "#548b54", color = NA) +
  geom_rect(aes(xmin = 0, ymin = n*2 - 0.7,
                xmax = percent, ymax = n*2 + 0.7), fill = "#548b54", color = NA) +
  geom_text(aes(-58, n*4, label = country_name), family = "DejaVu Sans Mono", color = "white", hjust = 0, size = 8.5, check_overlap = TRUE) +
  geom_text(aes(percent, n*2, label = paste(percent, "% ", sep = "")), family = "DejaVu Sans Mono", color = "white", hjust = 1, size = 4, check_overlap = TRUE) +
  annotate("text", 110, 17, label = "European\nRenewable Energy\nProduction", family = "Noto Sans Black", color = "black", hjust = 1, size = 11, lineheight = 0.9) +
  annotate("text", 12, -11.5, label = " Source: Eurostat Energy\nGraphic: Alex Norman", family = "Noto Sans Thin", color = "black", hjust = 0, size = 4, lineheight = 0.9) +
  scale_x_continuous(breaks = seq(0, 100, 25), labels = seq(0, 100, 25)) +
  theme_minimal(base_family = "DejaVu Sans Mono") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(family = "DejaVu Sans Mono", size = 12),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "grey", size = 0.1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.margin = margin(20, 20, 20, 20)
  )
  

ggsave(here::here("2020-32", "plots", "temp",
    paste0("energy-", format(Sys.time(), "%Y%m%d_%H%M%S"),
    ".png")), dpi = 320, width = 9, height = 9)

```